\section{Min-Cost Max-Flow}

\subsection{Problem}

Given a directed graph $G = (V, E)$ where each edge $(u, v)$ has a \textbf{capacity} $c(u,v)$ and a \textbf{cost} $w(u,v)$, find the maximum flow from source $s$ to sink $t$ that minimizes total cost.

$$\text{minimize} \sum_{(u,v) \in E} w(u,v) \cdot f(u,v) \quad \text{subject to} \quad \sum f = \text{max-flow}$$

The key idea: use \textbf{Shortest Path Augmentation} (SPFA / Bellman-Ford) to always push flow along the cheapest augmenting path.

\subsection{Example Graph}

\begin{diagram}[id="mcmf-graph", label="MCMF Network"]
\shape{G}{Graph}{
  nodes=["S","A","B","C","D","T"],
  edges=[("S","A",4),("S","B",3),("A","C",2),("A","D",3),("B","C",5),("B","D",1),("C","T",4),("D","T",3)],
  directed=true,
  show_weights=true,
  layout="hierarchical",
  orientation="LR"
}
\recolor{G.node[S]}{state=current}
\recolor{G.node[T]}{state=good}
\end{diagram}

Edge labels show capacity. Costs: $w(S,A)=1$, $w(S,B)=2$, $w(A,C)=3$, $w(A,D)=1$, $w(B,C)=1$, $w(B,D)=4$, $w(C,T)=2$, $w(D,T)=1$.

\subsection{Algorithm Walkthrough}

\begin{animation}[id="mcmf-spfa", label="MCMF via Shortest Path Augmentation"]
\shape{G}{Graph}{
  nodes=["S","A","B","C","D","T"],
  edges=[("S","A",4),("S","B",3),("A","C",2),("A","D",3),("B","C",5),("B","D",1),("C","T",4),("D","T",3)],
  directed=true,
  show_weights=true,
  layout="hierarchical",
  orientation="LR"
}
\shape{dist}{Array}{size=6, data=["$\\infty$","$\\infty$","$\\infty$","$\\infty$","$\\infty$","$\\infty$"], labels="0..5", label="dist[ ]"}
\shape{vars}{VariableWatch}{names=["flow","cost","path","bottleneck"], label="State"}

\step
\recolor{G.node[S]}{state=current}
\recolor{G.node[T]}{state=good}
\apply{vars.var[flow]}{value="0"}
\apply{vars.var[cost]}{value="0"}
\apply{vars.var[path]}{value="---"}
\apply{vars.var[bottleneck]}{value="---"}
\narrate{Network with source $S$ and sink $T$. Edge labels = capacity. We track total flow and cost.}

\step
\narrate{\textbf{Round 1:} Run SPFA from $S$ to find cheapest path. Initialize $dist[S] = 0$, all others $\infty$.}
\apply{dist.cell[0]}{value="0"}
\recolor{dist.cell[0]}{state=current}

\step
\narrate{Relax edges from $S$: $dist[A] = 0 + 1 = 1$, $dist[B] = 0 + 2 = 2$.}
\apply{dist.cell[1]}{value="1"}
\apply{dist.cell[2]}{value="2"}
\recolor{dist.cell[1]}{state=current}
\recolor{dist.cell[2]}{state=current}
\recolor{G.edge[(S,A)]}{state=current}
\recolor{G.edge[(S,B)]}{state=current}

\step
\narrate{Relax from $A$: $dist[D] = 1 + 1 = 2$ via $A \to D$ (cost 1). $dist[C] = 1 + 3 = 4$ via $A \to C$.}
\apply{dist.cell[3]}{value="4"}
\apply{dist.cell[4]}{value="2"}
\recolor{dist.cell[3]}{state=current}
\recolor{dist.cell[4]}{state=current}
\recolor{G.edge[(A,C)]}{state=current}
\recolor{G.edge[(A,D)]}{state=current}
\recolor{G.node[A]}{state=current}

\step
\narrate{Relax from $B$: $dist[C] = \min(4, 2+1) = 3$ improved via $B \to C$! $D$ not improved ($2+4=6 > 2$).}
\apply{dist.cell[3]}{value="3"}
\recolor{G.edge[(B,C)]}{state=good}
\recolor{G.node[B]}{state=current}

\step
\narrate{Relax from $D$: $dist[T] = 2 + 1 = 3$ via $D \to T$.}
\apply{dist.cell[5]}{value="3"}
\recolor{dist.cell[5]}{state=good}
\recolor{G.edge[(D,T)]}{state=current}

\step
\narrate{Relax from $C$: $dist[T] = \min(3, 3+2) = 3$ not improved. SPFA complete.}
\recolor{G.edge[(C,T)]}{state=dim}

\step
\narrate{Cheapest path: $S \to A \to D \to T$ with cost $1+1+1=3$. Bottleneck = $\min(4, 3, 3) = 3$.}
\recolor{G.edge[(S,A)]}{state=good}
\recolor{G.edge[(A,D)]}{state=good}
\recolor{G.edge[(D,T)]}{state=good}
\recolor{G.edge[(S,B)]}{state=dim}
\recolor{G.edge[(B,C)]}{state=dim}
\recolor{G.edge[(A,C)]}{state=dim}
\recolor{G.node[A]}{state=good}
\recolor{G.node[D]}{state=good}
\apply{vars.var[path]}{value="S-A-D-T"}
\apply{vars.var[bottleneck]}{value="3"}

\step
\narrate{Push 3 units along path. $flow = 3$, $cost = 3 \times 3 = 9$. Update residual capacities.}
\apply{vars.var[flow]}{value="3"}
\apply{vars.var[cost]}{value="9"}
\recolor{G.node[S]}{state=current}
\recolor{G.node[T]}{state=good}
\recolor{G.node[A]}{state=done}
\recolor{G.node[D]}{state=done}
\recolor{G.edge[(S,A)]}{state=done}
\recolor{G.edge[(A,D)]}{state=done}
\recolor{G.edge[(D,T)]}{state=done}
\recolor{G.edge[(S,B)]}{state=idle}
\recolor{G.edge[(B,C)]}{state=idle}
\recolor{G.edge[(A,C)]}{state=idle}

\step
\narrate{\textbf{Round 2:} Run SPFA again on residual graph. Reset distances.}
\apply{dist.cell[0]}{value="0"}
\apply{dist.cell[1]}{value="$\\infty$"}
\apply{dist.cell[2]}{value="$\\infty$"}
\apply{dist.cell[3]}{value="$\\infty$"}
\apply{dist.cell[4]}{value="$\\infty$"}
\apply{dist.cell[5]}{value="$\\infty$"}
\recolor{dist.cell[0]}{state=current}
\recolor{dist.cell[1]}{state=idle}
\recolor{dist.cell[2]}{state=idle}
\recolor{dist.cell[3]}{state=idle}
\recolor{dist.cell[4]}{state=idle}
\recolor{dist.cell[5]}{state=idle}
\recolor{G.node[A]}{state=idle}
\recolor{G.node[B]}{state=idle}
\recolor{G.node[D]}{state=idle}
\apply{vars.var[path]}{value="---"}
\apply{vars.var[bottleneck]}{value="---"}

\step
\narrate{SPFA finds: $S \to A \to C \to T$ (cost $1+3+2=6$, residual cap $\min(1, 2, 4)=1$) and $S \to B \to C \to T$ (cost $2+1+2=5$, cap $\min(3, 5, 4)=3$). Cheapest: $S \to B \to C \to T$.}
\apply{dist.cell[2]}{value="2"}
\apply{dist.cell[3]}{value="3"}
\apply{dist.cell[5]}{value="5"}
\recolor{dist.cell[2]}{state=current}
\recolor{dist.cell[3]}{state=current}
\recolor{dist.cell[5]}{state=good}

\step
\narrate{Path $S \to B \to C \to T$, cost $5$, bottleneck $= 3$. Push 3 units.}
\recolor{G.edge[(S,B)]}{state=good}
\recolor{G.edge[(B,C)]}{state=good}
\recolor{G.edge[(C,T)]}{state=good}
\recolor{G.node[B]}{state=good}
\recolor{G.node[C]}{state=good}
\apply{vars.var[path]}{value="S-B-C-T"}
\apply{vars.var[bottleneck]}{value="3"}

\step
\narrate{$flow = 3 + 3 = 6$, $cost = 9 + 3 \times 5 = 24$. Update residual.}
\apply{vars.var[flow]}{value="6"}
\apply{vars.var[cost]}{value="24"}
\recolor{G.edge[(S,B)]}{state=done}
\recolor{G.edge[(B,C)]}{state=done}
\recolor{G.edge[(C,T)]}{state=done}
\recolor{G.node[B]}{state=done}
\recolor{G.node[C]}{state=done}

\step
\narrate{\textbf{Round 3:} SPFA on residual. $S \to A$ has 1 residual cap. Find $S \to A \to C \to T$: cost $1+3+2=6$, cap $\min(1, 2, 1)=1$.}
\recolor{G.edge[(S,A)]}{state=good}
\recolor{G.edge[(A,C)]}{state=good}
\recolor{G.edge[(C,T)]}{state=good}
\recolor{G.node[A]}{state=good}
\recolor{G.node[C]}{state=good}
\apply{vars.var[path]}{value="S-A-C-T"}
\apply{vars.var[bottleneck]}{value="1"}

\step
\narrate{Push 1 unit. $flow = 7$, $cost = 24 + 1 \times 6 = 30$. No more $s$-$t$ paths in residual.}
\apply{vars.var[flow]}{value="7"}
\apply{vars.var[cost]}{value="30"}
\recolor{G.edge[(S,A)]}{state=done}
\recolor{G.edge[(A,C)]}{state=done}
\recolor{G.node[A]}{state=done}

\step
\narrate{\textbf{Result:} Max-flow $= 7$, Min-cost $= 30$. All source edges saturated: $S \to A$ used 4/4, $S \to B$ used 3/3.}
\recolor{G.node[S]}{state=done}
\recolor{G.node[T]}{state=done}
\apply{vars.var[path]}{value="done"}
\apply{vars.var[bottleneck]}{value="done"}

\end{animation}

\subsection{Complexity}

With SPFA (Bellman-Ford variant):

$$O(V \cdot E \cdot F)$$

where $F$ is the max flow value. Each SPFA round takes $O(VE)$ and pushes at least 1 unit of flow.

For competitive programming, the practical bound is usually fast enough when $|V| \le 500$ and $|E| \le 5000$.

\subsection{Implementation}

\begin{lstlisting}[language=Python]
from collections import deque

def mcmf(graph, s, t, n):
    """Min-Cost Max-Flow via SPFA augmentation."""
    total_flow = 0
    total_cost = 0

    while True:
        # SPFA to find cheapest s-t path
        dist = [float('inf')] * n
        dist[s] = 0
        in_queue = [False] * n
        parent = [-1] * n
        parent_edge = [-1] * n

        q = deque([s])
        in_queue[s] = True

        while q:
            u = q.popleft()
            in_queue[u] = False
            for i, (v, cap, cost, rev) in enumerate(graph[u]):
                if cap > 0 and dist[u] + cost < dist[v]:
                    dist[v] = dist[u] + cost
                    parent[v] = u
                    parent_edge[v] = i
                    if not in_queue[v]:
                        q.append(v)
                        in_queue[v] = True

        if dist[t] == float('inf'):
            break  # no augmenting path

        # Find bottleneck
        bottleneck = float('inf')
        v = t
        while v != s:
            u = parent[v]
            bottleneck = min(bottleneck, graph[u][parent_edge[v]][1])
            v = u

        # Augment
        v = t
        while v != s:
            u = parent[v]
            ei = parent_edge[v]
            graph[u][ei][1] -= bottleneck       # reduce cap
            ri = graph[u][ei][3]
            graph[v][ri][1] += bottleneck       # increase reverse cap
            v = u

        total_flow += bottleneck
        total_cost += bottleneck * dist[t]

    return total_flow, total_cost
\end{lstlisting}
