\begin{animation}[id="bfs-grid", label="BFS Shortest Path on Grid"]
\shape{g}{Grid}{rows=5, cols=5, data=[["S",".",".","X","."],[".","X",".",".","."],[".",".",".","X","."],[".","X","X",".","."],[".",".",".",".","G"]]}

\step
\recolor{g.cell[0][0]}{state=current}
\apply{g.cell[0][0]}{value=0}
\narrate{BFS from S(0,0) to G(4,4). X = wall. Numbers = distance from S.}

\step
\recolor{g.cell[0][0]}{state=done}
\recolor{g.cell[0][1]}{state=current}
\recolor{g.cell[1][0]}{state=current}
\apply{g.cell[0][1]}{value=1}
\apply{g.cell[1][0]}{value=1}
\narrate{Wave 1: neighbors of (0,0). Reach (0,1) and (1,0).}

\step
\recolor{g.cell[0][1]}{state=done}
\recolor{g.cell[1][0]}{state=done}
\recolor{g.cell[0][2]}{state=current}
\recolor{g.cell[2][0]}{state=current}
\apply{g.cell[0][2]}{value=2}
\apply{g.cell[2][0]}{value=2}
\narrate{Wave 2: reach (0,2) and (2,0). Wall at (1,1) blocks.}

\step
\recolor{g.cell[0][2]}{state=done}
\recolor{g.cell[2][0]}{state=done}
\recolor{g.cell[1][2]}{state=current}
\recolor{g.cell[2][1]}{state=current}
\recolor{g.cell[3][0]}{state=current}
\apply{g.cell[1][2]}{value=3}
\apply{g.cell[2][1]}{value=3}
\apply{g.cell[3][0]}{value=3}
\narrate{Wave 3: reach (1,2), (2,1), (3,0).}

\step
\recolor{g.cell[1][2]}{state=done}
\recolor{g.cell[2][1]}{state=done}
\recolor{g.cell[3][0]}{state=done}
\recolor{g.cell[1][3]}{state=current}
\recolor{g.cell[2][2]}{state=current}
\recolor{g.cell[4][0]}{state=current}
\apply{g.cell[1][3]}{value=4}
\apply{g.cell[2][2]}{value=4}
\apply{g.cell[4][0]}{value=4}
\narrate{Wave 4: reach (1,3), (2,2), (4,0). Wall (2,3) blocks east.}

\step
\recolor{g.cell[1][3]}{state=done}
\recolor{g.cell[2][2]}{state=done}
\recolor{g.cell[4][0]}{state=done}
\recolor{g.cell[1][4]}{state=current}
\recolor{g.cell[4][1]}{state=current}
\apply{g.cell[1][4]}{value=5}
\apply{g.cell[4][1]}{value=5}
\narrate{Wave 5: reach (1,4) and (4,1).}

\step
\recolor{g.cell[1][4]}{state=done}
\recolor{g.cell[4][1]}{state=done}
\recolor{g.cell[0][4]}{state=current}
\recolor{g.cell[2][4]}{state=current}
\recolor{g.cell[4][2]}{state=current}
\apply{g.cell[0][4]}{value=6}
\apply{g.cell[2][4]}{value=6}
\apply{g.cell[4][2]}{value=6}
\narrate{Wave 6: reach (0,4), (2,4), (4,2). Three new cells.}

\step
\recolor{g.cell[0][4]}{state=done}
\recolor{g.cell[2][4]}{state=done}
\recolor{g.cell[4][2]}{state=done}
\recolor{g.cell[3][4]}{state=current}
\recolor{g.cell[4][3]}{state=current}
\apply{g.cell[3][4]}{value=7}
\apply{g.cell[4][3]}{value=7}
\narrate{Wave 7: reach (3,4) and (4,3). Almost at goal.}

\step
\recolor{g.cell[3][4]}{state=done}
\recolor{g.cell[4][3]}{state=done}
\recolor{g.cell[3][3]}{state=done}
\recolor{g.cell[4][4]}{state=good}
\apply{g.cell[3][3]}{value=8}
\apply{g.cell[4][4]}{value=8}
\narrate{Wave 8: reached G(4,4) and (3,3). Shortest path = 8 steps.}
\end{animation}
